\texttt{davos}: The Python package smuggler

### 2022-04-06 paper/main.tex
updates to motivation

% --------------------------------------------- MOTIVATION & SIGNIFICANCE ----------------------------------------------
\section{Motivation and significance}
Modern scientific research often involves writing code for a variety of purposes throughout the research process, from designing experiments, to analyzing and visualizing data, to presenting findings via tutorials or public demos. One requirement common to nearly all research-related code it that its behavior and/or outputs must remain consistent and predictable every time it is run, both over time and across users. For example, this stability can be critical to ensuring that data are collected under constant conditions (e.g., across recordings, trials, or subjects) and that statistical analyses yield accurate, consistent results. Additionally, modern open science practices encourage sharing code and data publicly to enable others to explore, reproduce, learn from, and build upon existing work. Scientists, researchers, and educators may seek to share research-related code with a wide range of audiences including collaborators, students, the broader scientific community, and the general public.

Python has become one of the most widely used and fastest-growing scientific programming languages by combining an accessible, high-level syntax with the availability of powerful third-party tools that facilitate rapid development and collaboration \cite{MullEtal15}. The Python ecosystem offers an extensive data science toolkit, including platforms for interactive programming (e.g., Project Jupyter, \cite{KluyEtal16}; Google Colaboratory), community-maintained libraries for data manipulation (e.g., \texttt{NumPy} \cite{HarrEtal20}, \texttt{SciPy} \cite{VirtEtal20}, \texttt{Pandas} \cite{McKi10}) and visualization (e.g., \texttt{Matplotlib} \cite{Hunt07}, \texttt{seaborn} \cite{Wask21}), and myriad other tools. However, this also presents a challenge to writing and sharing stable, reproducible scientific Python code: different versions of the same third-party library may behave differently, use different syntax, add or drop support for specific functions or other libraries, address (or introduce) bugs, and so on. While these issues are present to some extent in any language or ecosystem, they have a particular impact on the Python community due to its unusually rapid growth relative to other languages. Ensuring consistent behavior over time and across users therefore typically requires ensuring that Python code is always run with a specific set of versions for each package used.

>>>>

% --------------------------------------------- MOTIVATION & SIGNIFICANCE ----------------------------------------------
\section{Motivation and significance}
Modern scientific research often involves writing code for a variety of purposes throughout the research process, from designing experiments, to analyzing and visualizing data, to presenting findings or techniques via tutorials or public demos. One requirement common to nearly all research-related code it that its behavior and/or outputs must remain consistent and predictable every time it is run, both over time and across users. For example, this stability can be critical to ensuring that data are collected under constant conditions (e.g., across recordings, trials, or subjects) and that statistical analyses yield accurate, reliable results. Additionally, modern open science practices encourage sharing code and data publicly to enable others to explore, reproduce, learn from, and build upon existing work. Scientists, researchers, and educators may seek to share research code with audiences that have a wide range of scientific and technical backgrounds, including collaborators, students, the broader scientific community, and the general public.

Python has become one of the most widely used and fastest-growing scientific programming languages by combining an accessible, high-level syntax with the availability of powerful third-party tools that facilitate rapid development and collaboration \cite{MullEtal15}. The Python ecosystem offers an extensive data science toolkit, with platforms for interactive programming (e.g., Project Jupyter \cite{KluyEtal16}, Google Colaboratory), community-maintained libraries for data manipulation (e.g., NumPy \cite{HarrEtal20}, SciPy \cite{VirtEtal20}, Pandas \cite{McKi10}) and visualization (e.g., Matplotlib \cite{Hunt07}, seaborn \cite{Wask21}), frameworks for training complex machine learning models (e.g., scikit-learn \cite{PedrEtal11}, TensorFlow \cite{AbadEtal15}, Hugging Face \cite{WolfEtal20}), and myriad other resources. However, this heavy emphasis on third-party libraries also presents a challenge to writing and sharing stable, reproducible scientific Python code: different versions of the same package may behave differently, use different syntax, add or drop support for specific functions or other libraries, address (or introduce) bugs, and so on. While these issues are present to some extent in any language or ecosystem, they have a particular impact on the Python community due to its unusually rapid growth relative to other languages. Ensuring Python code behaves consistently over time and across users therefore typically requires making sure it is always run with a specific set of versions for each package used.

One common approach to solving this problem is to create containerized or virtualized Python environments (e.g., using \href{https://www.docker.com/}{Docker}, \href{https://sylabs.io/singularity/}{Singularity}, or \href{https://docs.conda.io/en/latest/}{conda}) tailored to individual applications. Researchers can then share these environments alongside their code as configuration files that explicitly list required packages and versions, allowing other users to build identical copies for themselves. While effective, this approach bears two significant drawbacks: First it can substantially add to the technical knowledge, system resources, and setup required to share and run the encapsulated code.
